### 2023-04-12 header.tex

% **************GENERATED FILE, DO NOT EDIT**************

\title{My JuliaCon proceeding}

\author[1]{Taylor Allred}
\author[1]{Xinyi Li}
\author[1]{Ashton Wiersdorf}
\author[1]{Ben Greenman}
\author[1]{Ganesh Gopalakrishnan}
\affil[1]{University of Utah}

\keywords{Julia, Optimization, Game theory, Compiler}

\hypersetup{
pdftitle = {My JuliaCon proceeding},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, Optimization, Game theory, Compiler},
}