### 2023-08-17 header.tex
checkpoint: 1,2

% **************GENERATED FILE, DO NOT EDIT**************

\title{FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions}

\author[1]{Taylor Allred}
\author[1]{Xinyi Li}
\author[1]{Ashton Wiersdorf}
\author[1]{Ben Greenman}
\author[1]{Ganesh Gopalakrishnan}
\affil[1]{University of Utah}

\keywords{Julia, floating-point, debugging}

\hypersetup{
pdftitle = {FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Taylor Allred, Xinyi Li, Ashton Wiersdorf, Ben Greenman, Ganesh Gopalakrishnan},
pdfkeywords = {Julia, floating-point, debugging},
}