### 2023-08-17 paper.tex

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{microtype}

%% extra links:
%% - GPU Julia: https://cuda.juliagpu.org/stable/tutorials/introduction/
%% - GPU x2: https://nextjournal.com/sdanisch/julia-gpu-programming
%% - NaN vs missing https://discourse.julialang.org/t/is-there-any-reason-to-use-nan-instead-of-missing/84396

\begin{document}

% Misc typographical tweaks.
\setlength{\parindent}{10pt}
% \setmonofont{JuliaMono}[Extension = .ttf, Path = ./, UprightFont = *-Regular, ItalicFont = *-Italic]

\input{header}
\maketitle

\begin{abstract}
  Reliable numerical computations are central to scientific computing,
  but the floating-point arithmetic that enables large-scale models
  is fundamentally unreliable.
  Numeric exceptions are a common occurrence and can propagate through
  code, leading to flawed results.
  This paper presents FlowFPX, a toolkit for systematically debugging
  floating-point exceptions by recording their flow,
  coalescing exception contexts, and fuzzing in select locations.
  These tools help scientists discover when exceptions happen
  and track down their origin, smoothing the way to a reliable codebase.
\end{abstract}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\FloatTracker}{FloatTracker}
\newcommand{\FT}{\FloatTracker}
\newcommand{\Fp}{Floating-point} % hyphen or no?
\newcommand{\fp}{floating-point} % hyphen or no?
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zerocode}[1]{\zerowidth{\code{#1}}}

\section{Introduction}

Reliable numeric computations are central to high-performance computing,
machine learning, and scientific applications.
Yet the \fp{} arithmetic that powers these computations is fundamentally
unreliable~(\cref{s:background}).
Exceptional values, such as Not a Number (\Nan{}) and infinity (\Inf{}),
can and often do arise thanks to culprits such as roundoff error,
catastrophic cancellation, singular matrices, and vanishing
derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
Developers are responsible for guarding against exceptions, but this task
is difficult because many operations can generate and propagate exceptions.
Worst of all, operations such as less-than (\code{<}) can kill an exceptional value,
leaving no trace of the problem.
There is little tool support to assist in exception debugging,
thus (unsurprisingly) a quick GitHub search reports over 4,000 open issues
related to \NaN{} or \Inf{} exceptions~\cite{github-issues}.

This paper introduces \FlowFPX{}, a toolkit for debugging
\fp{} exceptions~(\cref{s:flowfpx})
that hase helped improve a variety of applications,
from ocean simulations to heat modes~(\cref{s:casestudies}):
\begin{itemize}
  \item
    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that
    selectively monitors for exceptions and fuzzes code for vulnerabilities.
  \item
    To visualize results, \FT{} adapts coalesced stack-trace graphs (CSTGs, or
    \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014}
    to summarize the program contexts that handled exceptional values.
  \item
    A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
\end{itemize}

\section{Floating-Point Exception Primer}
\label{s:background}

\begin{figure}[t]\centering
  \includegraphics[trim=0 0 0 60,clip,width=0.95\columnwidth]{fig/real_vs_fp.png}
  \caption{Floats are spread across the real number line.}
  \label{f:real-vs-fp}
\end{figure}

\Fp{} numbers use a finite number of bits to represent a spectrum of points along the real number line~(\cref{f:real-vs-fp}).
The implementation strategy is essentially scientific notation.
A \fp{} number packs a sign bit, an exponent, and a fraction part (also called the ``significand'' or ``mantissa'') into a bitstring (typically 64 or 32 bits, but 16-bit and 8-bit formats are on the rise~\cite{klowerLowprecisionClimateComputing2021,fp8}).
This representation supports very small and very large numbers in a narrow range of bits.

\begin{lstlisting}[language = Julia]
julia> prevfloat(typemax(Float64))
1.7976931348623157e308
julia> nextfloat(typemin(Float64))
-1.7976931348623157e308
\end{lstlisting}

The flip side is that most real numbers fall into the gaps between \fp{} numbers and must be rounded, which introduces error
and can lead to surprising results.
For example, adding the tiny Planck constant to the large Avogadro number results
in Avogadro's number after rounding:

\begin{lstlisting}[language = Julia]
julia> planck = 6.62607015e-34
6.62607015e-34
julia> avogadro = 6.02214076e23
6.02214076e23
julia> avogadro + planck == avogadro
true
\end{lstlisting}

This is an extreme example, but many operations on \fp{} numbers closer in magnitude induce a loss of accuracy.
Refer to the literature for more details, e.g.,~\cite{knuthArtComputerProgramming1997,torontoPracticallyAccurateFloatingPoint2014,mullerHandbookFloatingPointArithmetic2018}.

\subsection{Exceptions and Exceptional Values}
\label{s:exnvalue}

The IEEE~754 \fp{} standard~\cite{IEEEStandardBinary1985}
defines exceptions and exceptional values as the outcome of
operations that have ``no single universally acceptable result''~\cite{p-draft-1997}.
For example, dividing by zero and exceeding the \code{Float64} range
both lead to exceptions:

% julia> 2 / 0
% Inf
\begin{lstlisting}[language = Julia]
julia> 0 / 0
NaN
julia> avogadro^avogadro
Inf
julia> log(0)
-Inf
julia> Inf + NaN
NaN
\end{lstlisting}

IEEE~754 defers the question of how to handle such exceptions to application code.
It is up to developers to watch for \Nan{} and \Inf{} (and also subnormal numbers)
and implement an appropriate repair.
Without guards, bugs happen.

\subsection{Lifetime of an Unhandled Exceptions}
\label{s:to-kill-a-fp}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/genpropkill-outline.png}
  \caption{Gen, Prop, Kill: Lifetime of a \NaN{} value}
  \label{f:gpk}
\end{figure}

Exceptional values have a lifetime: they are born, or \emph{generated}, by some operation; they \emph{propagate} through other operations; and they either appear in the program output, go out of scope, or get \emph{killed} by a numeric operation.
\Cref{f:gpk} summarizes this \emph{gen}-\emph{prop}-\emph{kill} process.
The gens and props are straightforward; see above for examples~(\cref{s:exnvalue}).
The kills often arise from numeric comparisons (\code{<}, \code{=}, etc.),
but exponents (\code{1\^NaN}) and over-eager matrix optimizations can kill
exceptions~\cite{ddghlllprr-correctness-2022}.

To illustrate TODO

Consider the following two functions to find the maximum value in a list:

\begin{lstlisting}[language = Julia]
function max1(lst)
  max_seen = 0.0
  for x in lst
    # swap if x is not too small
    if ! (x <= max_seen)
      max_seen = x
    end
  end
  max_seen
end

function max2(lst)
  foldl(max, lst)
end
\end{lstlisting}

Both \code{max1} and \code{max2} compute the maximum in a list.
However, the \texttt{<=} operator used in the \code{max1} implementation does \emph{not} propagate \NaN{}s: \texttt{NaN <= 4} returns \texttt{false}, not \NaN{}.
Because of this, you might not get the result you expected:

\begin{lstlisting}[language = Julia]
max1([1, 5, NaN, 4])     # returns 4
max2([1, 5, NaN, 4])    # returns NaN
\end{lstlisting}

Not only is the result from \texttt{max1} problematic for obscuring the fact that there was a \NaN{} in the list, the result is also wrong!
The built-in \texttt{max} function correctly propagates \NaN{}s, so we can see from its return that we got a \NaN{} somewhere.
It's not entirely obvious \emph{where} that \NaN{} came from.

\section{\FlowFPX{}}
\label{s:flowfpx}

\FlowFPX{} is a set of tools to aid in tracking down \fp{} exceptions.
The two primary tools we cover in this paper are \FT{} and CSTG:

\begin{itemize}
\item \FT{} traces the lifetime of exceptional \fp{} values and enables inside-out fuzzing.
\item CSTG visualizes the tracing output from \FT{} in compact form.
\end{itemize}

\GPUFPX~\cite{llsflg-hpdc-2023} is another component of \FlowFPX{} that can track \fp{} exceptions inside a GPU kernel.
See the cited paper for more details on \GPUFPX.

%% TODO julia proposed consistent NaN handling,
%%  do we propagate the exact NaN
%%  https://github.com/JuliaLang/julia/issues/48523
%%
%% Ashton: @Ben we simply propagate whatever we got from the operation. If
%% Julia's behavior around an operator change, that should be preserved even
%% when using FloatTracker

\subsection{\FT{}}
\label{s:floattracker}

\FT{} is implemented as a Julia library, and is available for download on JuliaHub.\footnote{\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig/1.0.0}}

\subsubsection{Tracking Exceptional Values}
\label{s:trackingexceptionalvalues}

\FT{}'s first purpose is to monitor \fp{} operations for exceptional value lifetime events.
\FT{} interposes on arithmetic operations and standard library calls that involve a \fp{} value and logs a stack trace to that point whenever an exception is generated, propagated, or killed.
\FT{} can also log the arguments to the suspect call, so you can get \emph{normal, concrete} values to help discover what is going wrong.
A programmer can then use these stack traces to pinpoint where an exceptional value like \NaN{} came into being or where a guard against a \NaN{} kill is needed.

All the logs that \FT{} captures are straight from Julia's \texttt{stacktrace()} function.
\FT{} writes these logs to three files: one file for each kind of event in a \fp{} exception's lifetime.
This way, if a user of \FT{} is attempting to determine where a \NaN{} or an \Inf{} came from, they can look in the file for all the gen events.
Likewise, if the user wants to find where \NaN{}s get killed, then they can look in the kills file.

\subsubsection{Fuzzing}

The second role of \FT{} is to fuzz code from the inside out.
While can fuzz a program by giving it random inputs, this might not help you find \emph{all} the places where an exceptional value could crop up during a calculation.
Demmel et. al suggest fuzzing by injecting exceptional values at intermediate calculations;~\cite{ddghlllprr-correctness-2022} this is the approach that we enable with \FT{}.

\FT{}'s exception tracking infrastructure offers us a simple way to inject exceptional values wherever we want in our code: we are already interposing on \fp{} operations to monitor for exceptional lifetime values, so we can inject a \NaN{} at any of these points.

Since the injection happens at random, \FT{} also provides a way to save \NaN{} injections so that injections that exercise bugs can be replayed to examine the bug repeatedly and to verify that hardening efforts have indeed eliminated the vulnerability.

We have a global configuration system that allows us to turn on fuzzing and configure where, when, and how many \NaN{}s to inject, as well as whether or not to record or replay a recording:

\begin{itemize}
\item \texttt{replay::String} Name of a file to read an injection recording from. If set, this option makes all others irrelevant.
\item \texttt{record::String} Name of a file to write injection events to for later replay.
\item \texttt{active::Bool} Inject only if this is true.
\item \texttt{n\_inject::Int64} Inject at most this many \NaN{}s.
\item \texttt{odds::Int64} Inject if \texttt{rand(odds) == 1}.
\item \texttt{functions::Array} Inject if we are within the dynamic extent of any of these functions.
\item \texttt{libraries::Array} Inject if the top-most stack frame\footnote{After stripping off the stack frames that come from \FT{} itself.} comes from any of these libraries.
\end{itemize}

Then, each time we intercept a \fp{} operation, we check to see if we should inject a \NaN{}. First we check if we're replaying a recording: if we are, just do as the recording says. Otherwise, we check that the injector is \texttt{active}, we have \NaN{}s remaining to inject (\texttt{n\_inject}), and we roll an \texttt{odds}-sided die to see if we should inject. Finally, check the current stack trace against the lists of functions and libraries to inject in. If all looks good, we inject.

\subsubsection{\FT{} Internals}

While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, \FT{} works entirely in user land.
\FT{} provides custom types---\texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64}---that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
\FT{} also defines methods for every function in the \texttt{Base} module.
This lets \FT{} intercept all \fp{} operations involving \texttt{TrackedFloat} types.

\begin{lstlisting}[language = Julia]
function +(x::TrackedFloat64, y::TrackedFloat64)
  result = run_or_inject(+, x.val, y.val)
  check_error(+, result, x.val, y.val)
  TrackedFloat64(result)
end
\end{lstlisting}

This new method does a few things:

\begin{enumerate}
\item First it checks if it should inject a \NaN{} here.
\item If no \NaN{} is to be injected, it unwraps the \texttt{TrackedFloat}s to get to the primitive \texttt{Float} value, and computes the result normally.
\item It examines the arguments as well as the return value, and then uses \cref{fig:lifetime-class} to determine the type of event and logs it if it is of interest.
\item It wraps the result in a new \texttt{TrackedFloat} value.
\end{enumerate}

\begin{figure}[h]
  \centering
  \begin{tabular}{ccc}
    Arguments & Result & Event Type \\
    \hline
    No \NaN{} & \NaN{} & Gen \\
    \NaN{} & \NaN{} & Prop \\
    \NaN{} & No \NaN{} & Kill \\
  \end{tabular}
  \label{fig:lifetime-class}
  \caption{Exceptional value event classification rules}
\end{figure}

The same classification rules apply to \Inf{} and would apply to any other value of interest.

These new methods had to be defined for every combination of tracked/untracked arguments;
fortunately Julia's metaprogramming abilities made it simple for us to scale without the code becoming unwieldy.
We made a macro that first loops through all the \texttt{TrackedFloat} types that we wished to implement, and then looped through all of the operators that we wished to support.
This approached saved us over 10x the lines a non-macro approach would have required.\footnote{We implemented overrides for 645 function/type combinations. At roughly 5 lines per variant, this is 3225 lines of code. Our implementation at time of writing weighs in at 218 lines of code, which is \emph{including} whitespace and comments.}
This approach was inspired by \texttt{Sherlogs.jl}.\cite{kMilanklSherlogsJl2021}

\subsection{CSTG}
\label{s:cstg}

\FT{} can produce copious amounts of log files which can be challenging to sift through manually.
Coalesced Stack Trace Graphs, or CSTGs~\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
CSTG is a stand-alone tool and was not designed with \FT{} specifically in mind.
However, it pairs well with \FT{} and makes analyzing the log files easier.

CSTG takes a set of stack traces and makes a graph: for each stack trace, it creates a node for every frame, and then creates or increments an edge between two adjacent frames, as seen in \cref{fig:cstg_demo}.

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{./fig/cstg_static_diagram.png}
  \caption{CSTG makes a graph by connecting adjacent stack frames}
  \label{fig:cstg_demo}
\end{figure}

These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
Higher nodes in the graph show function calls progressively closer to the top-level call that spawned them.
Thick lines and high numbers indicate flows with the highest volume leading to errors---though thin areas are still worth investigating.
CSTG provides a map to aid in contextualizing a stack trace in a log file.

\subsection{\GPUFPX{}}
\label{s:gpufpx}

% HELP: @ben, I'm not sure what paper you mean by "Julia CPU"
Many programs offload work onto GPUs, which are no less susceptible to \fp{} exceptions than CPUs.
In some regards they are \emph{more} susceptible because GPUs do not have the same exception-handling mechanisms that CPUs do.
Thus, tools to catch \fp{} exceptions on the GPU are essential.
\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}~\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
\GPUFPX{} can detect floating-point exceptions on the GPU that GPU code cannot do on its own.
\GPUFPX{} reports these exceptions back to the CPU for further analysis.

\FT{} and \GPUFPX{} can cover the entire computational stack, as seen in \cref{fig:ft_and_gpu_fpx}.
Work remains to tighten the integration between \FT{} and \GPUFPX{} to reduce manual work to employ both tools for heterogeneous computation.

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{./fig/float_tracker_context_boxes.png}
  \caption{FloatTracker and GPU-FPX compliment each other}
  \label{fig:ft_and_gpu_fpx}
\end{figure}

\section{Case Studies}
\label{s:casestudies}

In this section we'll discuss some case studies where tools from \FlowFPX{} aided in finding or fixing \fp{} issues.
\Cref{s:sw} covers an ocean simulation that produces \NaN{}s in the output---a common and frustrating occurrence for many a scientist. \Cref{s:ode} puts the popular \texttt{OrdinaryDiffEq} solver library under a fuzzer to harden it against bad behavior in the presence of exceptional \fp{} values. \Cref{s:finch} fuzzes another library---this time a library for specifying partial differential equations (PDEs). Finally, in \cref{s:ocean}, we examine another ocean simulation library---this time the library runs on the GPU. This puts it out of reach of \FT{}, but \GPUFPX{} can handle it. In addition to the case studies that we did ourselves, \cref{s:safari} is an ``in the wild'' usage of \FT{} that helped researchers uncover the source of a \NaN{} in their own code without help from us.

\subsection{ShallowWaters}
\label{s:sw}

\texttt{ShallowWaters}~\cite{klowerNumberFormatsError2020,klowerPositsAlternativeFloats2019,klowerLowprecisionClimateComputing2021} simulates the flow of water over a seabed.
This is not a general shallow waters simulation library---it is designed primarily to showcase numerical methods involving 16-bit floats.

The Courant-Friedrichs-Lewy (``CFL'') parameter roughly describes the size of the time step to take in running the simulation.
Normally this value sits between 0 and 1---smaller values mean the simulation runs slower but with higher accuracy, whereas higher values mean the simulation takes bigger time steps and consequently runs faster, but results might be inaccurate.
The primary reason for problems arising comes from information not having enough time to propagate through the simulation before the next time step is taken.

It is not uncommon to encounter cases where setting the CFL parameter too high results in \NaN{}s in the result.
Driving the CFL parameter low enough to excise the \NaN{}s can make the simulation too long and unwieldy to be comfortable to use.
Identifying where \NaN{}s begin to arise can help point the way to where some mathematical refinement can be used to make the simulation both fast and accurate enough for practical use.

We dialed the CFL parameter up to extraordinarily high values and we started seeing \NaN{}s in the result.
\code{ShallowWaters} detected the instability, but not before producing some bad output.

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{./fig/shallow_waters_cfl_diff.png}
  \caption{Effect of high CFL parameters: on the left, normal output. On the right: \NaN{}s cause white holes and gaps in the result.}
  \label{fig:sw_nans}
\end{figure}

The white regions in the right graph of \cref{fig:sw_nans} are regions where \NaN{}s crept into the computation and destroyed the results.
In one regard, this is not as bad as a \NaN{} kill---knowing that there are \NaN{}s in the computation means that we have a chance to fix it, whereas a \NaN{} kill might silently give us the wrong answer.
Incidentally, we know that there were no \NaN{} kills in this test all the events in our kill logs are from calling the \texttt{isnan} function---a clear sign that the program is deliberately taking action on the presence of a \NaN{}.

Finding the sources of \NaN{}s can be tricky.
We expected \NaN{}s in this case because we set the CFL parameter up so high, but tracking down the sources of \NaN{}s generally is a difficult and time-consuming problem.

\FT{} solves this problem for us by automatically tracking all sources of potential \NaN{} generation.
And in this case \texttt{ShallowWaters} makes this doubly easy:
\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of \fp{} precision, and allows you to specify a \fp{} type to use in the computation.
We simply asked it to use our \texttt{TrackedFloat32} type, and then the entire simulation ran under the supervision of FloatTracker.

\begin{lstlisting}[language = Julia]
run_model(T=TrackedFloat32,
          cfl=10, Ndays=100, nx=100, L_ratio=1,
          bc="nonperiodic",
          wind_forcing_x="double_gyre",
          topography="seamount")
\end{lstlisting}

This produced almost 17000 lines of logs in the gens file. Below is a sample stack trace from these logs: (formatted for clarity)

\begin{verbatim}
-([-Inf, -Inf])   FT/TrackedFloat.jl:106
momentum_u!       SW/rhs.jl:246
rhs_nonlinear!    SW/rhs.jl:50
rhs!              SW/rhs.jl:14 [inlined]
time_integration  SW/time_integration.jl:77
run_model         SW/run_model.jl:37
top-level
\end{verbatim}

We can see that the \NaN{} appeared as the result of $-\infty - -\infty$ (first line) and that subtraction happened in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
Thus, with little effort we were able to track down where the \NaN{}s were originating.

Now the question is: where did the $\infty$ come from?
We can look at the file of the $\infty$ gens where we see the following: (formatted for clarity)

\begin{verbatim}
^([-1.515f31, 2])          FT/TrackedFloat.jl:138
literal_pow()              intfuncs.jl:325
…
materialize(^)             broadcast.jl:860
top-level getproperty(...) examples/sw_nan_tf.jl:14
…
\end{verbatim}

It looks like the $-\infty$ came from $-1.515e31^2$---no wonder a Float32 type couldn't handle something that large.
Now we have some concrete values to work with: instead of wondering where an opaque \Inf{} or a \NaN{} came from, we have \emph{specific} values that a domain expert can use to diagnose the problem.

\subsubsection{CSTGs Help Us Get a Bigger Picture}

Running FloatTracker can produce a lot of logs.
To get a quick overview of the most common paths to problems, we can use CSTG to get an overview of what stack frames appear the most often.

We can produce a CSTG of both the \Inf{} as well as the \NaN{} logs.
\Cref{fig:sw_nan_cstg} and \cref{fig:sw_inf_cstg} show what the two graphs look like.\footnote{As with the call stack logs, these have been edited for clarity. We only remove a few boxes in between—mostly due to space constraints in this paper.}

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{fig/sw_nan_cstg_clean.png}
  \caption{CSTG of \NaN{} generation logs from ShallowWaters.}
  \label{fig:sw_nan_cstg}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{fig/sw_inf_cstg_clean.png}
  \caption{CSTG of \Inf{} generation logs from the same run as~\cref{fig:sw_nan_cstg}.}
  \label{fig:sw_inf_cstg}
\end{figure}

We can see in \cref{fig:sw_nan_cstg} that a thick line goes from the top-level call down to a call to \texttt{+} goes down the right side of the graph.
This indicates that the majority of the flows took this path.
That means we might be able to fix the majority of our problems by inspecting that particular flow.
\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{\^}---a promising place to start looking for improvements.

\subsubsection{CSTG Diffing}

Another one of CSTGs abilities is \emph{graph diffing}.
We took the \FT{} logs from running ShallowWaters and split the logs at the 10\% mark.
We used CSTG to diff the flow graphs between the first 10\% of the logs and the remaining 90\%.
The diff can be seen in \cref{fig:cstg_diff_demo}.

\begin{figure}[t]
  \centering
  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
  \caption{Diff between the early and latter stages of ShallowWaters}
  \label{fig:cstg_diff_demo}
\end{figure}

The green lines indicate flows that were \emph{new} in the latter 90\% of the logs.
The red lines indicate flows that were \emph{only in} the first 10\% of the logs.
What this means depends on the particular program; a domain expert can use these clues to e.g. find where instability started in the first part of the program, and where that instability got reinforced later on.
In this case, we can see that \texttt{momentum\_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.

\subsection{OrdinaryDiffEq}
\label{s:ode}

To exercise the fuzzing abilities of \FT{}, we tried fuzzing \texttt{NBodySimulator}\footnote{\url{https://github.com/SciML/NBodySimulator.jl}} from the SciML team, which simulates the gravitational interactions between two or more massive bodies.
However, when we used \FT{}, we got no \fp{} operations.
We discovered that \texttt{NBodySimulator} merely sets up the system of equations to solve, then hands all the work of running the simulation off to the \texttt{OrdinaryDiffEq}\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl}} library from the same team.

We configured \FT{} to inject a single \NaN{} randomly within the scope of the \texttt{NBodySimulator} and \texttt{OrdinaryDiffEq} libraries.
We found a run where something curious happened: the library reported that it had detected a \NaN{} and would then exit.
However, after printing that message, the program went into an infinite loop.

Using CSTGs to guide our search, we found a \NaN{} kill that manifested repeatedly in the logs.
The stack traces for that kill originated from inside the \texttt{solve.jl} file in the \texttt{OrdinaryDiffEq} library.

\begin{verbatim}
<([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
solve!           at ODE/solve.jl:515
…
\end{verbatim}

The relevant part of \texttt{solve.jl} looks like this:

% file here: ~/Research/ode_debug/dev/OrdinaryDiffEq/src/solve.jl
% see line 514

\begin{lstlisting}[language = Julia]
while !isempty(time_stops)
  while tdir * t < first(time_stops)
    # do integration work
    # pop_off(time_stops)
  end
end
\end{lstlisting}

The problem here is on line 2 when \texttt{tdir} is \NaN{}: multiplication propagates \NaN{}s, and comparison kills it, so the condition on the inner \texttt{while} loop is \emph{always} false.
No work got done in the inner loop, and so the size of the \texttt{time\_stops} vector never decreases.

Our kill logs led us right to this line and let us know which of the arguments to \texttt{<} was a \NaN{}.
This is a perfect example of how \NaN{} kills can influence control flow to go awry.
In this case, the problem was apparent besides what our logs told us and it manifested as an infinite loop.
More dangerous cases can occur when the influence of a bad comparison is not as readily observable.
In any case, \FT{} can help monitor for kills and help fuzz to find kills before they arise in production.

We were able to use the injection recording to reproduce the bug in the GitHub issue that we opened.\footnote{\url{https://github.com/SciML/OrdinaryDiffEq.jl/issues/1939}}

\subsection{Finch}
\label{s:finch}

Finch is a domain-specific language for specifying
PDEs~\cite{heislerFinchDomainSpecific2022}.\footnote{Not to be confused
with the Finch loop optimizer~\cite{adka-cgo-2023}.}
In the spirit of FEniCS~\cite{fenics} and related
tools~\cite{freefem,openfoam,dune,firedrake},
Finch helps scientists quickly convert math into code.
What sets Finch apart is its flexibility.
It supports multiple discretization methods (finite element and finite
volume) and multiple backends (Julia, \CPP{}, \Dendro{}~\cite{dendro}).
Furthermore it strives to output code that humans can easily fine-tune.

Fuzzing with \FT{} revealed two places where Finch needed
protection against user input.
The first was when reading an input mesh.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/16}{2023-06-06}}
A \NaN{} injected in the mesh led to a crash further on:

\begin{verbatim}
  BoundsError: attempt to access 1-element
   Vector{Int64} at index [2]
\end{verbatim}

Defensive programming solves the issue.
The second place was in setting bounds for the
solver.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/17}{2023-06-06}}
Here, a \NaN{} could leave bounds uninitialized, leading to a bounds error.
Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
appear in unstable systems.
%% advection2d, high cfl and high T (time) led to NaN during simulation

\subsection{\GPUFPX{}: Oceananigans}
\label{s:ocean}

Oceananigans~\cite{OceananigansJOSS} is simulation package for incompressible
fluid dynamics that can generate code for Nvidia GPUs.
For example, the following program (from the project readme) simulates turbulance:

\begin{lstlisting}[language = Julia]
using Oceananigans
grid = RectilinearGrid(GPU(),
  size=(128, 128), x=(0, 2π), y=(0, 2π),
  topology=(Periodic, Periodic, Flat))
model = NonhydrostaticModel(; grid,
  advection=WENO())
ϵ(x, y, z) = 2rand() - 1
set!(model, u=ϵ, v=ϵ)
simulation = Simulation(model;
  Δt=0.01, stop_time=4)
run!(simulation)
\end{lstlisting}

\GPUFPX{} provides detailed feedback on this program.
The output in~\cref{f:gpufpx} shows that 21 kernels
appear and generate six floating-point exceptions.
There are three \NaN{}s, one \Inf{}, and two division
by zero errors.
The report is a starting point for further investigation
of the reliability of the example.
%% TODO exactly how?

\begin{figure}[t]\centering
  \begin{tabular}[t]{ll}
    \begin{tabular}[t]{lr}
      \zerocode{-{}- FP64 Operations -{}-} \\
      \code{Total NaN:} & \code{2} \\
      \code{Total INF:} & \code{1} \\
      \code{Total subnormal:} & \code{0} \\
      \code{Total div0:} & \code{2} \\
    \end{tabular}
    &
    \begin{tabular}[t]{lr}
      \zerocode{-{}- FP32 Operations -{}-} \\
      \code{Total NaN:} & \code{1} \\
      \code{Total INF:} & \code{0} \\
      \code{Total subnormal:} & \code{0} \\
      \code{Total div0:} & \code{0} \\
    \end{tabular}
  \end{tabular}
  \begin{tabular}{lr}
    \zerocode{-{}- Other Stats -{}-} \\
    \code{Kernels:} & \code{21}
  \end{tabular}

  \caption{Example \GPUFPX{} output}
  \label{f:gpufpx}
\end{figure}

\subsection{RxInfer: No Help Needed}
\label{s:safari}

\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.

\section{Related Work}
\label{s:related}

\subsection{Floating-Point Error Analysis}

Demmel et al.\cite{ddghlllprr-correctness-2022} examine \fp{} exceptional value handling in the BLAS and LAPACK libraries and propose solutions to inconsistencies.
They describe a proposal to pack numeric codes into an \texttt{INFO\_ARRAY} that mirrors our gen-prop-kill framework.
They also suggest a fuzzing technique that we realize in \FT{}:
instead of fuzzing the \emph{inputs} to a program, we fuzz by randomly injecting exceptional values at \emph{any} point where there is an arithmetic operation. Demmel suggests that this approach is useful since any \fp{} operation could lead to an exceptional value if the inputs are in dangerous ranges.

Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} develop a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
They include error visualization and work on how to avoid error-magnifying ``badlands''.
They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on the exceptional values themselves.

In a similar vein, Herbie\cite{panchekhaAutomaticallyImprovingAccuracy2015} is a tool for automatically rewriting arithmetic expressions to reduce \fp{} error.
\FT{} tracks down exceptional \fp{} values; once the location of a gen is found, Herbie could be employed to rewrite the arithmetic expression, if \fp{} error is the cause of the exceptional value.
The Odyssey\cite{misbackOdysseyInteractiveWorkbench2023} workbench provides an interactive interface to Herbie.

\subsection{Diagnosing floating-point exceptions}

The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired the use of a custom number type to intercept operations on a number.
In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.

% TODO: make sure that what I said about what FPSpy can track is true.
FPSpy~\cite{dindaSpyingFloatingPoint2020} is an \texttt{LD\_PRELOAD} shared library that works on unmodified x86 binaries.
It can monitor a program during execution for any \fp{} arithmetic issues, such as division by zero, denormalization, underflow, and overflow.
\FT{} in contrast focuses on the lifetime of exceptional values; \FT{} can notice when a \NaN{} gets killed, for instance, while that is not an event that FPSpy appears to be able to track.
FPSpy has the advantage of being lightweight enough to run on production code for certain loads.

\subsection{Stack Trace Graphs}

CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
STAT is a tool for collecting, analyzing, and visualizing stack traces from thousands of concurrent processes, and is designed to highlight anomalies quickly.
It produces visualizations similar to those of CSTG, thought CSTG offers more compact views.

\subsection{Floating-Point Exceptions on the GPU}

FPChecker~\cite{l-ase-2019} is a tool to report \fp{} exceptions occurring on the GPU.
FPChecker relies on LLVM-level instrumentation of GPU kernels, and so cannot run on the plethora of closed-source GPU kernels in usage today.

BinFPE~\cite{llg-soap-2022} is another tool in the same space;
BinFPE performs SASS-level analysis of GPU kernels, but is limited in that it is slow and does not catch errors that alter control flow.
The latter deficiency is particularly worrisome, as we have seen, silent \NaN{} kills can invalidate results without the user noticing.

\GPUFPX{}~\cite{llsflg-hpdc-2023} improves on the work of FPChecker and BinFPE by being more performant and catching a wider set of errors, including those that alter control flow.
\GPUFPX{} is a shared library that, like FPSpy uses \texttt{LD\_PRELOAD} to work on unmodified binaries.
\GPUFPX{} runs on CUDA cores from NVIDIA and reports total numbers of exceptional values.
Like \FT{}, \GPUFPX{} can catch \NaN{} gens and kills, but it does this on the GPU where \FT{} doesn't apply.
Despite these improvements, \GPUFPX{} is limited by the closed-source nature of common GPU cores, and cannot report at the rich level of source detail that \FT{} can.

Work is ongoing to use \GPUFPX{} and \FT{} in tandem.

% Error analysis, typical answer.
% Difficult.
% CITE

% % Herbie, improving accuracy automatically.
% % ParetoHerbie

% %% http://eprints.maths.manchester.ac.uk/2873/1/fami22.pdf
% %% https://drive.google.com/drive/folders/1HkgMpxi6hsqLSZj9tSnuCxnFExPPjmvx?usp=sharing
% %% https://docs.google.com/presentation/d/1ueqzb5QY9YHmtY2DfY_hpN9oufSOcXErLWsJOZQxgzo/edit?usp=sharing
% %% https://ieeexplore.ieee.org/document/8916392
% %% https://link.springer.com/chapter/10.1007/978-3-319-73721-8_24

% FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
% instrumentation"~\cite{ltlg-iiswc-2022}
% % FPChecker has a nice table of tools; some already mentioned here

% 6-8,24,25
% Arnab Das scalable yet rigorous
% Marc Daumas Certification of Bounds Expres
% David Delmas Towards an Industrial Use of
% Alexey Solovyev FPTaylor Results table
% Titolo absint

% GPU-FPX state of art, companion project~\cite{llsflg-hpdc-2023}.
% Latest in line of work.
% BinFPE exn checking, slower than GPU-FPX, 
% \cite{llg-soap-2022}.
% FPChecker, early pioneer, fpx in gpus, llvm-level instrumentation
% if compiled with Clang, but most gpus use closed NVCC compiler~\cite{l-ase-2019}.

\section{Discussion}
\label{s:discussion}
% future work

\subsection{Performance}
\label{s:discussion-performance}

% TODO: citations!!!!
% TODO: real numbers!!!
\FT{} incurs significant overhead on the order of 100x slower than a non-instrumented run of the same program.
To put this number into context, Valgrind runs with a similar level of slowdown.

In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.

When injection is turned on we defer the stack checks as late as we can.
Once the fuzzer runs out of \NaN{}s, it effectively turns off, so the rest of the fuzzing session proceeds much more quickly.
That said, gathering stack traces still incurs a heavy overhead during logging when interesting events occur;
we recommend limiting the number and kinds of events logged for optimal performance.

\subsection{Tracking More than \NaN{}s and \Inf{}s}

FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
For example, \FT{} could be made to monitor extreme values: in some circumstances it would be useful to notice when a value started climbing above or falling below a certain threshold.

We spent most of the time engineering and testing \NaN{} tracking and injection.
Adding \Inf{} tracking took a few hours.
Adding other values of interest presents no fundamental difficulties.

\subsection{Enhanced Fuzzing}

While fuzzing is useful for discovering issues, its success rate
is low because \emph{every} \fp{} operation is a candidate
for injection.
Even operations that are already well-defended against \NaN{}s are candidates.
\FT{} could use two sorts of tools for improving injection.
First, fine-grained control to let users decide where not to inject.
Second, tools for understanding the context of an injection point
after the fact.
Program slicing is especially relevant to the latter point and effectively
what we did by hand when fuzzing Finch (\cref{s:finch}).
For each operation, an expert needs to study the values that feed into it to
decide whether they are protected or not.

%% TODO example, start here at glbvertex https://github.com/paralab/Finch/blob/master/src/grid.jl#L360

\subsection{Tracking Exceptions in External Libraries}

\FT{}'s scope is limited to Julia code: it cannot track the lifetime of exceptional values that are born, live, or die in external libraries, such as C programs.
One approach would be to create an API for \FT{} so that programs using these external libraries could leverage \FT{}'s tooling to monitor the interface between Julia and e.g. external C programs.

\subsection{Future Work}

\FlowFPX{} is a growing project and we welcome feedback from domain experts
on how to best suit their workflow.
If you try any of our tools, please send us comments, suggestions, and bug reports via the tools' repositories on GitHub.

\section{Acknowledgments}

We would like to thank Alex Larsen and Rob Durst for their help with reading drafts of the paper and feedback on the presentation.
We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.
Finally, many thanks to Milan Klöwer, the author of the \texttt{Sherlogs} library, for inspiration on the architecture of \FT{}.

\input{bib.tex}

\end{document}

% Below is some Emacs-specific stuff to help me (Ashton) write this paper. If
% you're using Emacs, you should try using the new jinx package for
% spell-checking!

% Local Variables:
% jinx-local-words: "BinFPE CFL CSTG's CSTGs Courant Demmel FPChecker Friedrichs Herbie JuliaHub Klöwer Lewy OrdinaryDiffEq RxInfer ShallowWaters Sherlogs TrackedFloat isnan"
% citar-bibliography: ("./ref.bib")
% End: